Under-approximation of reachability through static analysis in multivalued asynchronous networks

### 2013-04-01 parts/def-ph.tex
section 2.4

\subsection{Consequences of the restrictions}
%Given these hypotheses, we can denote: $\pfp_s(a) = a_i$ the local steady-state of $a$ in state $s$.
We denote by $\update(s)$ the state in which all cooperative sorts are updated.
\begin{align*}
  \update(s) = s \Cap \{ \pfp_{s}(a) \mid a \in CS \} \enspace.
\end{align*}

\pref{th:update} states that from any state, there exists a scenario updating the cooperative sorts of this state.
\begin{theorem}
\label{th:update}
  $\forall s \in \PHl, \exists \delta \in \restriction{\Sce}{1}(s), s \PHplay \delta = \update(s)$
\end{theorem}
\begin{proof}
  Let $a$ be a cooperative sort so that $\PHget{s}{a} \neq \pfp_s(a)$.
  Given the definition of $\pfp_s(a)$, there exists a scenario updating $a$ in $s$.
  As there is no cycle in between the cooperative sorts and an updated cooperative sort cannot evolve, at most $|\cs|$ updates can be done.
\end{proof}
\pref{th:hcompcomp} states that for a given state $s$ and for any action $h = \PHhit{a_i}{b_j}{b_k}$ between well-formed components, $h$ can be played after a series of hits that do not prevent it to be fired.
\pref{th:hcscomp} states the same if $a$ is a well-formed cooperative sort, under the condition that $a_i$ reflects the state of the neighborhood of $a$.
\begin{theorem}
\label{th:hcompcomp}
  $\forall s \in \PHl, \forall h = \PHhit{a_i}{b_j}{b_k} \in \PHh, a,b \in \components,$\\
  $(\PHget{s}{a} = a_i \wedge \PHget{s}{b} = b_j) \Rightarrow (\exists \delta \in \restriction{\Sce}{1}(s), \text{$h$ can be played in $s \PHplay \delta$})$
\end{theorem}
\begin{proof}
  From \pref{th:update}, there exists a scenario $\delta$ so that: $(s \PHplay \delta) = \update(s)$.
  As $a,b \in \components$, $a_i \in s \PHplay \delta$ and $b_j \in s \PHplay \delta$.
  Finally, by definition of $\update(s)$, no action of priority $1$ can be played in $s \PHplay \delta$, thus $h$ can be played in $s \PHplay \delta$.
\end{proof}
\begin{theorem}
\label{th:hcscomp}
  $\forall s \in \PHl, \forall h = \PHhit{a_i}{b_j}{b_k} \in \PHh, a \in \cs, b \in \components$\\
  $(\PHget{s}{a} = a_i \wedge \PHget{s}{b} = b_j \wedge \pfp_s(a) = a_i) \Rightarrow (\exists \delta \in \restriction{\Sce}{1}(s), \text{$h$ can be played in $s \PHplay \delta$})$
\end{theorem}
\begin{proof}
  From \pref{th:update}, there exists a scenario $\delta$ so that: $(s \PHplay \delta) = \update(s)$.
  As $a_i \in \pfp_s(a)$, $a_i \in s \PHplay \delta$.
  Furthermore, $b \in \components$, thus $b_j \in s \PHplay \delta$.
  Finally, by definition of $\update(s)$, no action of priority $1$ can be played in $s \PHplay \delta$, thus $h$ can be played in $s \PHplay \delta$.
\end{proof}

>>>>

\subsection{Consequences of the restrictions}

The previous restrictions lead to a particular class of PH models that can be efficiently studied.
In this subsection, we give several general theorems that can be derived from these restrictions,
and which will help building the static analysis of \pref{sec:sa}.

We first denote by $\update(s)$ the state equivalent to $s$ but in which all cooperative sorts are updated (\pref{def:update}).
This state is unique due to the properties of $\pfp$ given in the previous subsection.
Then, \pref{th:update} states that from any state, there exists a scenario updating the cooperative sorts of this state.
%
\begin{definition}[$\update : \PHl \rightarrow \PHl$]
\label{def:update}
  For all $s \in \PHl$, we define:
  \begin{align*}
    \update(s) = s \Cap \{ \pfp_{s}(a) \mid a \in CS \} \enspace.
  \end{align*}
\end{definition}
%
\begin{theorem}
\label{th:update}
  $\forall s \in \PHl, \exists \delta \in \restriction{\Sce}{1}(s), s \PHplay \delta = \update(s)$
\end{theorem}
\begin{proof}
  Let $a$ be a cooperative sort so that $\PHget{s}{a} \neq \pfp_s(a)$.
  Given the definition of $\pfp_s(a)$, there exists a scenario $\delta$ updating $a$ in $s$ so that
  no scenario in $\restriction{\Sce}{1}(s \PHplay \delta)$ can change the active process of $a$.
  As there is no cycle of actions between the cooperative sorts (\pref{cr:cyclefreeness}) and an updated cooperative sort cannot evolve, at most $|\cs|$ updates can be done.
\end{proof}
\pref{th:hcompcomp} states that for a given state $s$ and for any action $h = \PHhit{a_i}{b_j}{b_k}$ between components,
if $\PHget{s}{a} = a_i$ and $\PHget{s}{b} = b_j$, then
$h$ can always be played after a series of hits (and these hits do not prevent it to be fired).
\pref{th:hcscomp} states the same if $a$ is a cooperative sort, under the condition that $a$ is updated in $s$.
%$a_i$ reflects the state of the neighborhood of $a$.
\begin{theorem}
\label{th:hcompcomp}
  $\forall s \in \PHl, \forall a,b \in \components, \forall h = \PHhit{a_i}{b_j}{b_k} \in \PHh,$\\
  $(\PHget{s}{a} = a_i \wedge \PHget{s}{b} = b_j) \Rightarrow (\exists \delta \in \restriction{\Sce}{1}(s), \text{$h$ can be played in $s \PHplay \delta$})$
\end{theorem}
\begin{proof}
  From \pref{th:update}, there exists such a scenario $\delta$ so that: $(s \PHplay \delta) = \update(s)$.
  As $a,b \in \components$, $a_i \in s \PHplay \delta$ and $b_j \in s \PHplay \delta$.
  Finally, by definition of $\update(s)$, no action of priority $1$ can be played in $s \PHplay \delta$, thus $h$ can be played in $s \PHplay \delta$.
\end{proof}
\begin{theorem}
\label{th:hcscomp}
  $\forall s \in \PHl, \forall h = \PHhit{a_i}{b_j}{b_k} \in \PHh, a \in \cs, b \in \components$\\
  $(\PHget{s}{a} = a_i \wedge \PHget{s}{b} = b_j \wedge \pfp_s(a) = a_i) \Rightarrow (\exists \delta \in \restriction{\Sce}{1}(s), \text{$h$ can be played in $s \PHplay \delta$})$
\end{theorem}
\begin{proof}
  Similar to the proof of \pref{th:hcompcomp}, and
  %From \pref{th:update}, there exists such a scenario $\delta$ so that: $(s \PHplay \delta) = \update(s)$.
  as $a_i \in \pfp_s(a)$, $a_i \in s \PHplay \delta$.
  %Furthermore, $b \in \components$, thus $b_j \in s \PHplay \delta$.
  %Finally, by definition of $\update(s)$, no action of priority $1$ can be played in $s \PHplay \delta$, thus $h$ can be played in $s \PHplay \delta$.
\end{proof}
